---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

# nomisr

[![License: MIT](https://img.shields.io/badge/License-MIT-blue.svg)](https://opensource.org/licenses/MIT)
[![CRAN_Status_Badge](https://www.r-pkg.org/badges/version/nomisr)](https://cran.r-project.org/package=nomisr)
[![GitHub tag](https://img.shields.io/github/tag/evanodell/nomisr.svg)](https://github.com/evanodell/nomisr)
[![](https://cranlogs.r-pkg.org/badges/grand-total/nomisr)](https://dgrtwo.shinyapps.io/cranview/)
[![Travis-CI Build Status](https://travis-ci.org/evanodell/nomisr.svg?branch=master)](https://travis-ci.org/evanodell/nomisr)
[![AppVeyor Build Status](https://ci.appveyor.com/api/projects/status/github/evanodell/nomisr?branch=master&svg=true)](https://ci.appveyor.com/project/evanodell/nomisr)
[![Coverage Status](https://img.shields.io/codecov/c/github/evanodell/nomisr/master.svg)](https://codecov.io/github/evanodell/nomisr?branch=master)
[![DOI](https://zenodo.org/badge/118144805.svg)](https://zenodo.org/badge/latestdoi/118144805)

`nomisr` is for accessing [Nomis](https://www.nomisweb.co.uk/) data with R. Nomis contains data from the Census, the Labour Force Survey, DWP benefit statistics and other economic and demographic data, and is maintained on behalf of the Office for National Statistics by the University of Durham.

The `nomisr` package provides functions to find what data is available, the  variables and query options for different datasets and a function for downloading data. `nomisr` returns data in [tibble](https://cran.r-project.org/package=tibble) format. Most of the data available through `nomisr` is based around statistical geographies, with some a handful of exceptions.

## Installation

You can install `nomisr` from github with:

```{r gh-installation, eval = FALSE}
# install.packages("devtools")
devtools::install_github("evanodell/nomisr")
```

## Using `nomisr`

The example below gets the latest data on Jobseeker's Allowance with rates and proportions, on a national level, with all male claimants and workforce.

```{r example}
 library(nomisr)
 z <- nomis_get_data(id = "NM_1_1", time = "latest", geography = "TYPE499", measures=c(20100, 20201), sex=5)
 tibble::glimpse(z)
```

There is a lot of data available through Nomis, and there are some limits to the amount of data that can be retrieved within a certain period of time, although those are not published. For more details, see the [full API documentation](https://www.nomisweb.co.uk/api/v01/help) from Nomis. Full package documentation is available at [docs.evanodell.com/nomisr](https://docs.evanodell.com/nomisr)


## Meta

Please note that this project is released with a [Contributor Code of Conduct](CONDUCT.md). By participating in this project you agree to abide by its terms.

Please note that this project is not affiliated with the Office for National Statistics or the University of Durham.

License: MIT


