---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

# nomisr

[![License: MIT](https://img.shields.io/badge/License-MIT-blue.svg)](https://opensource.org/licenses/MIT)
[![CRAN_Status_Badge](https://www.r-pkg.org/badges/version/nomisr)](https://cran.r-project.org/package=nomisr)
[![GitHub tag](https://img.shields.io/github/tag/evanodell/nomisr.svg)](https://github.com/evanodell/nomisr)
[![](https://cranlogs.r-pkg.org/badges/grand-total/nomisr)](https://dgrtwo.shinyapps.io/cranview/)
[![Travis-CI Build Status](https://travis-ci.org/evanodell/nomisr.svg?branch=master)](https://travis-ci.org/evanodell/nomisr)
[![AppVeyor Build Status](https://ci.appveyor.com/api/projects/status/github/evanodell/nomisr?branch=master&svg=true)](https://ci.appveyor.com/project/evanodell/nomisr)
[![Coverage Status](https://img.shields.io/codecov/c/github/evanodell/nomisr/master.svg)](https://codecov.io/github/evanodell/nomisr?branch=master)
[![DOI](https://zenodo.org/badge/118144805.svg)](https://zenodo.org/badge/latestdoi/118144805)

`nomisr` is for accessing [Nomis](https://www.nomisweb.co.uk/) data with R. Nomis contains labour market statistics, data on benefits, demographic data and census data.

## Installation

You can install nomisr from github with:

```{r gh-installation, eval = FALSE}
# install.packages("devtools")
devtools::install_github("evanodell/nomisr")
```

## Using `nomisr`

The example below gets the latest data on Jobseeker's Allowance with rates and proportions, on a national level, with all male claimants and workforce.

```{r example}
 library(nomisr)
 z <- nomis_get_data(id="NM_1_1", time="latest", geography="TYPE499", measures=c(20100, 20201), sex=5)
 tibble::glimpse(z)
```

There is a lot of data available through Nomis, and there are some limits to the amount of data that can be retrieved within a certain period of time, although those are not published. For more details, see the [full API documentation](https://www.nomisweb.co.uk/api/v01/help) from Nomis. Full package documentation is available at [docs.evanodell.com/nomisr](https://docs.evanodell.com/nomisr)

