---
title: "Introduction to nomisr"
author: "Evan Odell"
date: "`r Sys.Date()`"
output: 
    html_document:
        toc: true
        toc_float: true
        theme: flatly
vignette: >
  %\VignetteIndexEntry{Introduction to nomisr}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Introduction to the Nomis API

Nomis provides free access to up-to-date official statistics from the Office for National Statistics (ONS), and is operated by the University of Durham.

There is a lot of data available through Nomis, and there are some limits to the amount of data that can be retrieved within a certain period of time, although those are not published. For more details, see the [full API documentation](https://www.nomisweb.co.uk/api/v01/help) from Nomis.

Nomis data is based around administrative and statistical geographies, and a particular geography should be specified when downloading data.

# Accessing Nomis data

`nomisr` is designed around a pipeline of three key functions: `nomis_data_info()`, `nomis_codes()` and `nomis_get_data()`.


## Querying data availability

Use the `nomis_data_info()` function without any parameters to get a tibble 
with metadata for all available datasets:

```{r, echo=FALSE}
  library(nomisr) 
  x <- nomis_data_info()
  head(x)
```


`nomis_data_info()` can also be used to query metadata from a specific dataset, using its ID. The example below uses the "LC4408EW - Tenure by number of persons per bedroom in household by household type" dataset from the 2011 census, including the 3 nested data frames including metadata important to making specific queries.

```{r, echo=FALSE}
  y <- nomis_data_info("NM_893_1")
  tibble::glimpse(y)
  
  y$annotations.annotation
  
  y$components.attribute
  
  y$components.dimension
```


When a tibble with metadata for all datasets or a specific dataset is returned, three of the columns: `annotations.annotation`, `components.attribute` and `components.dimension` are lists of nested data frames. `annotations.annotation` contains metadata on the dataset, including units and current status. `components.attribute` contains more detailed status metadata. `components.dimension` contains the grouping and summary variables available in the dataset, which vary between different datasets.

## Downloading data

Vast amounts of data are available through Nomis and so to avoid overwhelming the API, it is good practice to query what variables are available, using `nomis_codes`. This queries the different kinds of geography codes available through the API for the "LC4408EW - Tenure bynumber of persons per bedroom in household by household type" dataset.

```{r, echo=FALSE}
  a <- nomis_codes(id = "NM_893_1", code = "geography", type = "type")
  a

```


Using the information above, we can now query the latest data on bedroom occupancy per household type in different NHS clinical commissioning groups.

```{r, echo=FALSE}
 z <- nomis_get_data(id = "NM_893_1", time = "latest", geography = "TYPE266")
 tibble::glimpse(z)
```



