---
title: "Introduction to nomisr"
author: "Evan Odell"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```


## Querying data availability

Use the `nomis_data_info()` function without any parameters to get a tibble 
with metadata for all available datasets:

```{r, echo=FALSE, results='asis'}
  library(nomisr) 
  x <- nomis_data_info()
  head(x)
```


`nomis_data_info()` can also be used to query metadata from a specific 
dataset, using its ID. The example below uses the "UV035 - Distance travelled 
to work" dataset.

```{r, echo=FALSE, results='asis'}
  y <- nomis_data_info("NM_5_1")
  head(y)
```



When a tibble with metadata for all datasets or a specific dataset is 
returned, three of the columns: `annotations.annotation`, 
`components.attribute` and `components.dimension` are lists of nested 
data frames. `annotations.annotation` contains metadata on the dataset, 
including units and current status. `components.attribute` contains more 
detailed status metadata. `components.dimension` contains the grouping and 
summary variables available in the dataset, which vary between different 
datasets.



## Downloading data

A stand

```{r, echo=FALSE, results='asis'}
 library(nomisr)
 z <- nomis_get_data(id="NM_1_1", time="latest", geography="TYPE499", measures=c(20100, 20201), sex=5)
 tibble::glimpse(z)
```



