---
title: "Introduction to nomisr"
author: "Evan Odell"
date: "`r Sys.Date()`"
output: 
    html_document:
        toc: true
        toc_float: true
        theme: flatly
vignette: >
  %\VignetteIndexEntry{Introduction to nomisr}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Introduction to the Nomis API

Nomis provides free access to up-to-date official statistics from the Office for National Statistics (ONS), and is operated by the University of Durham.

There is a lot of data available through Nomis, and there are some limits to the amount of data that can be retrieved within a certain period of time, although those are not published. For more details, see the [full API documentation](https://www.nomisweb.co.uk/api/v01/help) from Nomis.

Nomis data is based around administrative and statistical geographies, and a particular geography should be specified when downloading data.

# Accessing Nomis data

`nomisr` is designed around a pipeline of three key functions: `nomis_data_info()`, `nomis_codes()` and `nomis_get_data()`.

## Querying data availability

Use the `nomis_data_info()` function without any parameters to get a tibble 
with metadata for all available datasets:

```{r, echo=TRUE}
library(nomisr)
x <- nomis_data_info()
head(x)
```


`nomis_data_info()` can also be used to query metadata from a specific dataset, using its ID. The example below uses the "LC4408EW - Tenure by number of persons per bedroom in household by household type" dataset from the 2011 census, including the 3 nested data frames including metadata important to making specific queries.

```{r, echo=TRUE}
library(dplyr, warn.conflicts = F)

y <- nomis_data_info("NM_893_1")
y$annotations.annotation %>% class()
#> [1] "list"
y$annotations.annotation[[1]] %>% class()
#> [1] "data.frame"
y %>% pull(annotations.annotation) %>% class()
#> [1] "list"
y %>% pull(annotations.annotation)%>% .[[1]] %>% class()
#> [1] "data.frame"
y %>% pull(annotations.annotation) %>% purrr::pluck(1) %>% class()
#> [1] "data.frame"
y %>% pull(annotations.annotation) %>% purrr::map(1) %>% class()
#> [1] "list"
y %>% tidyr::unnest(annotations.annotation) %>% glimpse()
```


When a tibble with metadata for all datasets or a specific dataset is returned, three of the columns: `annotations.annotation`, `components.attribute` and `components.dimension` are lists of nested data frames. `annotations.annotation` contains metadata on the dataset, including units and current status. `components.attribute` contains more detailed status metadata. `components.dimension` contains the grouping and summary variables available in the dataset, which vary between different datasets.

### Searching for data

`nomisr` also contains the `nomis_search()` function to search for datasets on particular topics. By default, it searches for a given string in the names and descriptions of available datasets.

```{r, echo=TRUE}

a <- nomis_search('claimants')

```

Nomis datasets also contain keywords, which `nomis_search()` will return metadata for if `keywords = TRUE`.


```{r, echo=TRUE}

b <- nomis_search('Claimants', keywords = TRUE)

```


## Querying data variables


Vast amounts of data are available through Nomis and so to avoid overwhelming the API, it is good practice to query what variables are available, using `nomis_codes()`. This queries the different kinds of geography codes available through the API for the "LC4408EW - Tenure bynumber of persons per bedroom in household by household type" dataset.

```{r, echo=TRUE}
  a <- nomis_codes(id = "NM_893_1", code = "geography", type = "type")

  a
```


The vast majority (98% as of February 2018) of Nomis datasets include 

## Downloading data

Using the information above, we can now query the latest data on bedroom occupancy per household type in different NHS clinical commissioning groups.

```{r, echo=TRUE}
 z <- nomis_get_data(id = "NM_893_1", time = "latest", geography = "TYPE266")
 tibble::glimpse(z)
```






